---
title: "Problem Set template"
author: "Fernanda Sobrino"
date: "29/03/2020"
output: 
  pdf_document:
    number_sections: yes
  html_document:
    df_print: paged
urlcolor: blue
---
```{r message=FALSE, warning=FALSE}
library(tidyverse)
Sys.setlocale(locale = 'USA')
library(tidyverse)
library(dslabs)
```

<!-- .Rmd files use  markdown, a text mark up language, to provide formating.--> 
<!--Text include within these strange arrows are comments and will not show up when you knit-->

# Front matter
This submission is my work alone and complies with the 30535 integrity policy.

Add your initials to indicate your agreement: **__**

Add your collaborators: **__**

Late coins used this pset: 0. Late coins left: X. 
<!--You may use up to two for a given assignment.)-->



# Exercises

##  1.1 explore gdp growth:

### a

```{r message=FALSE, warning=FALSE}
data("gapminder")

df = gapminder

df %>% arrange(country, year)-> df

df %>% group_by(country) %>% summarise(laggdp = lag(gdp)) -> lag_gdp

df$lag_gdp = lag_gdp$laggdp
```


### b


```{r message=FALSE, warning=FALSE}
df$gdp_growth = df$gdp -df$lag_gdp
df %>% group_by(continent,year)  %>% summarise(growth = mean(gdp_growth, na.rm = T)) ->grt_df  
grt_df %>% ggplot(aes(year, growth, color = continent)) + 
  geom_line() +labs(title = 'average annual growth by continent')
```

Oceania and Africa's growth is close to zero, others' change largely overtime.

## 1.2



```{r message=FALSE, warning=FALSE}
dfc = df[, c('year', 'country', 'continent', 'gdp', 'lag_gdp')] %>% na.omit()
dfc %>% group_by(country)  %>%
  summarise(overall_rate =  log(last(lag_gdp)) - first(log(lag_gdp))) ->cty_corss

left_join(cty_corss, unique(df[,c('country', 'continent')])) -> cc
mean_df = cc %>% group_by(continent) %>% summarise(mean = mean(overall_rate))
mean_df =left_join(cc, mean_df)
mean_df$prop  = mean_df$overall_rate / mean_df$mean
mean_df %>% arrange(prop) %>% head(10) -> less_dvlp
less_dvlp
```


## 1.3

```{r message=FALSE, warning=FALSE}
df$growth_rate = log(df$gdp) - log(df$lag_gdp) 
df %>% group_by(continent,year)  %>%
  summarise(ct_rate = mean(growth_rate, na.rm = T)) ->cr_df  
df %>% group_by(country,year)  %>%
  summarise(cty_rate = mean(growth_rate, na.rm = T)) %>% na.omit() ->ctr_df 
left_join(ctr_df, unique(df[,c('country', 'continent')])) -> cc
left_join(cc, cr_df) -> comp_df
comp_df$above = comp_df$cty_rate > comp_df$ct_rate
comp_df %>% group_by(country) %>% summarise(n_above = sum(above)) -> abv_df
abv_df %>% arrange(n_above %>% desc()) %>% head(10) # with some countries have same number as chile

```



## 2.1 Exploring variation

###  2.1.1a

```{r message=FALSE, warning=FALSE}
df %>% ggplot(aes(infant_mortality)) + geom_freqpoly()



```

The distribution is right skewed. Not so surprised, such negative things typical has such a pattern.


### 2.1.1b

```{r message=FALSE, warning=FALSE}
df %>% ggplot(aes(infant_mortality)) + geom_freqpoly(binwidth  = 100)

df %>% ggplot(aes(infant_mortality)) + geom_freqpoly(binwidth  = 0.01)
```

When binwidth goes to large, the skewness can be small.

### 2.1.2a

```{r message=FALSE, warning=FALSE}
life_cty = pivot_wider(df, id_cols = year, names_from  = country, values_from   = life_expectancy)
life_cty$mean = life_cty[,-1] %>% apply(MARGIN = 1, mean)
life_cty %>% ggplot(aes(x = year, y = mean)) + geom_line() + labs(title = 'increase in life expectancy')
df %>% ggplot(aes(life_expectancy)) + geom_freqpoly(binwidth  = 0.5)
df %>% ggplot(aes(life_expectancy)) + geom_freqpoly(binwidth  = 5)

```

It is left skewed.



### 2.1.2 b

```{r message=FALSE, warning=FALSE}

(life_cty[,-1] < 40 )%>% sum()
(life_cty[,-1] > 80 )%>% sum()
(life_cty[,-1] ) %>% max() ->mxy
lc = c(0, 0)
(apply((life_cty[,-1]) , 2, max) == mxy) %>% which() %>% names() ->nm
life_cty[, nm] %>% unlist() %>% which.max() ->yrid
life_cty$year[yrid]  ->yr
nm
yr
```

198 country-year pairs report a life expectancy under 40,  334 over 80? Hong Kong, China with the highest life expectancy.

### 2.1.3

if you leave binwidth unset, using `bins = 30`,  set xlim to half the range , so only half a bar shows.

### 2.2.a

- bar chart: NA is considered as  category.
- In a histogram, NA is ignored #
- value in histogram  has order.




### 2.2.b

Ignore NA by computing


# Exploratory Data Analysis

## 1

### a

```{r message=FALSE, warning=FALSE}
data("diamonds")


lm(price ~ ., diamonds) %>% summary()
```

carat and clarity are most important numeric and categorical variables, respectively.


### b

```{r message=FALSE, warning=FALSE}
cor.test(diamonds$carat, diamonds$price)
```
It is positively correlated, correlation coefficient is 0.92.

### c


It ignore the interaction between variables.


## 2

### a

```{r message=FALSE, warning=FALSE}
dfa = diamonds %>% count(color, cut) %>% group_by(color) %>%
  mutate(prop = n / sum(n))
dfa %>% head()


```


### b

```{r message=FALSE, warning=FALSE}
dfb = diamonds %>% count(cut, color) %>% group_by(cut) %>%
  mutate(prop = n / sum(n)) 
dfb %>% head()
```


### c

```{r message=FALSE, warning=FALSE}
ggplot(dfb, aes( color, prop)) +
  geom_bar(stat="identity",position='dodge') +
  facet_wrap(~cut)
```


### c



## 3

### a

```{r message=FALSE, warning=FALSE}
p = diamonds %>% ggplot() +
  geom_freqpoly(mapping = aes(x = price,
                              color = cut_width(carat, 0.4)), bins = 20)
p
p = diamonds %>% ggplot() +
  geom_freqpoly(mapping = aes(x = price,
                              color = cut_number(carat, 15)), bins = 20)
p
```


### b

It affect both skewness and kurtosis.


## 4

```{r message=FALSE, warning=FALSE}
diamonds %>% ggplot() +
  geom_density(mapping = aes(x = carat,
                             color = cut_width(price, 2500))) + labs(title = 'distribution of caret with price group as reference.')

```


## 5

```{r}
diamonds %>% ggplot +
  geom_boxplot(mapping = aes(x = cut_number(carat, 10),
                             y = price)) 
```


The range and variation of large diamonds' price is larger than small ones,it is surprised me. Maybe that small diamond has some relative practical usage, for larger diamond, the usage is much more vague, so that the price is also larger, and vague, the pricing is hard or unstable, so that much more variable.

## 6

```{r message=FALSE, warning=FALSE}
diamonds %>% ggplot() +
  geom_violin(mapping = aes(x = cut, y = price,
                            color = cut_number(carat, 3)))
```


## 7



```{r message=FALSE, warning=FALSE}
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = x, y = y)) +
  coord_cartesian(xlim = c(4, 11), ylim = c(4, 11))

ggplot(data = diamonds) +
  geom_bin2d(mapping = aes(x = x, y = y)) +
  coord_cartesian(xlim = c(4, 11), ylim = c(4, 11))
```

I find the  scatter plot is better, it is simple, the second one is complex, i prefer to sort a table  than looking at this plot to find outliers.









